---
title: "molly_OriginalHomeworkCode_03"
author: "Yinuo Mao"
date: "2025-02-11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
zombies <- read.csv("zombies.csv")
```

1. Calculate the population mean and standard deviation for each quantitative random variable (height, weight, age, number of zombies killed, and years of education). NOTE: You will not want to use the built in var() and sd() commands as these are for samples.
```{r}
# Height
height_mean <- mean(zombies$height)
height_sd <- sqrt(mean((zombies$height - height_mean)^2))

# Weight
weight_mean <- mean(zombies$weight)
weight_sd <- sqrt(mean((zombies$weight - weight_mean)^2))

# Age
age_mean <- mean(zombies$age)
age_sd <- sqrt(mean((zombies$age - age_mean)^2))

# Number of Zombies Killed
zombies_killed_mean <- mean(zombies$zombies_killed)
zombies_killed_sd <- sqrt(mean((zombies$zombies_killed - zombies_killed_mean)^2))

# Years of Education
education_mean <- mean(zombies$years_of_education)
education_sd <- sqrt(mean((zombies$years_of_education - education_mean)^2))

# Display Results
pop_stats <- data.frame(Variable = c("Height", "Weight", "Age", "Zombies Killed", "Years of Education"), Mean = c(height_mean, weight_mean, age_mean, zombies_killed_mean, education_mean), SD = c(height_sd, weight_sd, age_sd, zombies_killed_sd, education_sd))

pop_stats
```

2. Use {ggplot} to make boxplots of each of these variables by gender.
```{r}
library(ggplot2)
# Height by Gender
ggplot(zombies, aes(x = gender, y = height)) + geom_boxplot() + ggtitle("Height by Gender")

# Weight by Gender
ggplot(zombies, aes(x = gender, y = weight)) + geom_boxplot() + ggtitle("Weight by Gender")

# Age by Gender
ggplot(zombies, aes(x = gender, y = age)) + geom_boxplot() + ggtitle("Age by Gender")

# Zombies Killed by Gender
ggplot(zombies, aes(x = gender, y = zombies_killed)) + geom_boxplot() + ggtitle("Zombies Killed by Gender")

# Years of Education by Gender
ggplot(zombies, aes(x = gender, y = years_of_education)) + geom_boxplot() + ggtitle("Years of Education by Gender")
```

3. Use {ggplot} to make scatterplots of height and weight in relation to age. Do these variables seem to be related? In what way?
```{r}
# Height vs Age
ggplot(zombies, aes(x = age, y = height)) + geom_point() + geom_smooth(method = 'lm') + ggtitle("Height vs Age")

# Weight vs Age
ggplot(zombies, aes(x = age, y = weight)) + geom_point() + geom_smooth(method = 'lm') + ggtitle("Weight vs Age")

## Height and age have a stronger positive relationship than that of weight and age.
```

4. Using histograms and Q-Q plots, check whether the quantitative variables seem to be drawn from a normal distribution. Which seem to be and which do not (hint: not all are drawn from the normal distribution)? For those that are not normal, can you determine from which common distribution they are drawn?
```{r}
par(mfrow = c(2, 2))

# Height
hist(zombies$height, main = "Height", xlab = "Height")
qqnorm(zombies$height); qqline(zombies$height)

# Weight
hist(zombies$weight, main = "Weight", xlab = "Weight")
qqnorm(zombies$weight); qqline(zombies$weight)

# Age
hist(zombies$age, main = "Age", xlab = "Age")
qqnorm(zombies$age); qqline(zombies$age)

# Zombies Killed
hist(zombies$zombies_killed, main = "Zombies Killed", xlab = "Zombies Killed")
qqnorm(zombies$zombies_killed); qqline(zombies$zombies_killed)

# Years of Education
hist(zombies$years_of_education, main = "Years of Education", xlab = "Years of Education")
qqnorm(zombies$years_of_education); qqline(zombies$years_of_education)
par(mfrow = c(1, 1))

## Hight: the variable is a normal distribution. The histogram is symmetric and bell-shaped, and the Q-Q plots closely align with the line.
## Weight: the variable is a normal distribution. The histogram is closely symmetric and bell-shaped, and the Q-Q plots closely align with the line.
## Age: the variable is roughly normal. The histogram is symmetric and bell-shaped, and the Q-Q plot shows minor deviations from the line.
## Zombies Killed: the variable is not normal. The histogram is right-skewed, and the Q-Q plot shows a deviation from the diagonal line. This should be the Poisson distribution.
## Years of Education: this variable is not normal. The histogram is right-skewed, and the Q-Q plot shows a strong deviation from the diagonal line. This should be the discrete distribution.
```


5. Now use the sample() function to sample ONE subset of 30 zombie survivors (without replacement) from this population and calculate the mean and sample standard deviation for each variable. Also estimate the standard error for each variable, and construct the 95% confidence interval for each mean. Note that for the variables that are not drawn from the normal distribution, you may need to base your estimate of the CIs on slightly different code than for the normal…
```{r}
# Take a sample of 30 zombies without replacement
set.seed(123)
zombie_sample <- zombies[sample(nrow(zombies), 30, replace = FALSE), ]

# Calculate sample statistics for each variable
variables <- c("height", "weight", "age", "zombies_killed", "years_of_education")

for(var in variables)
    # Calculate mean
    sample_mean <- mean(zombie_sample[[var]])
    
    # Calculate sample sd
    sample_sd <- sd(zombie_sample[[var]])
    
    # Calculate se
    se <- sample_sd/sqrt(30)
    
    # Calculate 95% ci
    ci <- c(sample_mean - qt(0.975, 29)*se, 
            sample_mean + qt(0.975, 29)*se)
    
    # Print results
    cat("\nResults for", var)
    cat("Sample Mean:", round(sample_mean, 3))
    cat("Sample SD:", round(sample_sd, 3))
    cat("Standard Error:", round(se, 3))
    cat("95% CI:", round(ci, 3))
```

6. Now draw 99 more random samples of 30 zombie apocalypse survivors, and calculate the mean for each variable for each of these samples. Together with the first sample you drew, you now have a set of 100 means for each variable (each based on 30 observations), which constitutes a sampling distribution for each variable. What are the means and standard deviations of this distribution of means for each variable? How do the standard deviations of means compare to the standard errors estimated in [5]? What do these sampling distributions look like (a graph might help here)? Are they normally distributed? What about for those variables that you concluded were not originally drawn from a normal distribution?
```{r}
# Set up a data frame to store means from all 100 samples
variables <- c("age", "height", "weight", "iq", "years_of_education", "num_children", "optimism")
all_sample_means <- matrix(nrow = 100, ncol = length(variables))
colnames(all_sample_means) <- variables

# For reproducibility
set.seed(123)

# Draw 100 samples and calculate means for each variable
for(i in 1:100) {
  zombie_sample <- zombies[sample(nrow(zombies), 30, replace = FALSE), ]
  for(j in 1:length(variables)) {
    all_sample_means[i, j] <- mean(zombie_sample[[variables[j]]])
  }
}

# Convert to data
all_sample_means <- as.data.frame(all_sample_means)

# Create results table for sampling distributions
sampling_results <- data.frame(
  Variable = variables,
  Mean_of_Means = numeric(length(variables)),
  SD_of_Means = numeric(length(variables)),
  Theoretical_SE = numeric(length(variables))
)

# Calculate statistics for the sampling distributions
for(i in 1:length(variables)) {
  var <- variables[i]
  sampling_results$Mean_of_Means[i] <- mean(all_sample_means[[var]])
  sampling_results$SD_of_Means[i] <- sd(all_sample_means[[var]])
  sampling_results$Theoretical_SE[i] <- sd(zombies[[var]])/sqrt(30)
}

# Print results
print(sampling_results)

# Create histograms for each variable (with explicit breaks)
plot_sampling_dist <- function() {
  # Set up plotting area with 2 rows and 4 columns
  par(mfrow = c(2, 4))
  
  # Loop through variables and create histograms with explicit breaks
  for(var in variables) {
    # Get the data for this variable
    data_var <- all_sample_means[[var]]
    
    # Check for valid data
    if(length(data_var) > 0 && !all(is.na(data_var)) && var(data_var) > 0) {
      # Create histogram with fixed number of breaks
      hist(data_var, 
           breaks = 10,  # Fixed number of breaks
           main = paste("Sampling Distribution of", var),
           xlab = "Sample Mean",
           probability = TRUE,
           col = "lightblue",
           border = "black")
      
      # Add a density curve
      x <- seq(min(data_var), max(data_var), length.out = 100)
      y <- dnorm(x, mean = mean(data_var), sd = sd(data_var))
      lines(x, y, col = "red", lwd = 2)
    } else {
      # Create an empty plot with a message if data is invalid
      plot(0, 0, type = "n", axes = FALSE, 
           main = paste("Cannot plot", var), 
           xlab = "", ylab = "")
      text(0, 0, "Invalid data for histogram")
    }
  }
  
  # Reset to one plot per figure
  par(mfrow = c(1, 1))
}

# Create Q-Q plots for each variable's sampling distribution
plot_qq_sampling <- function() {
  par(mfrow = c(2, 4))
  for(var in variables) {
    # Get the data for this variable
    data_var <- all_sample_means[[var]]
    
    # Check for valid data
    if(length(data_var) > 0 && !all(is.na(data_var)) && var(data_var) > 0) {
      qqnorm(data_var, main = paste("Q-Q Plot for", var, "Means"))
      qqline(data_var, col = "red")
    } else {
      # Create an empty plot with a message if data is invalid
      plot(0, 0, type = "n", axes = FALSE, 
           main = paste("Cannot plot Q-Q for", var), 
           xlab = "", ylab = "")
      text(0, 0, "Invalid data for Q-Q plot")
    }
  }
  par(mfrow = c(1, 1))
}

# Display plots
plot_sampling_dist()
plot_qq_sampling()

# Test for normality
normality_results <- data.frame(
  Variable = variables,
  P_Value = numeric(length(variables)),
  Normal = character(length(variables))
)

for(i in 1:length(variables)) {
  var <- variables[i]
  data_var <- all_sample_means[[var]]
  
  # Check for valid data
  if(length(data_var) > 0 && !all(is.na(data_var)) && var(data_var) > 0) {
    test <- shapiro.test(data_var)
    normality_results$P_Value[i] <- test$p.value
    normality_results$Normal[i] <- ifelse(test$p.value > 0.05, "Yes", "No")
  } else {
    normality_results$P_Value[i] <- NA
    normality_results$Normal[i] <- "Could not test"
  }
}

# Print normality test results
print(normality_results)
```
